---
title: "voi.remote for Value of Information calculation - based on Chris Jackson's VoI package"
author: "Eric Yu, William Xu"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
 \usepackage[utf8]{inputenc}
 %\VignetteIndexEntry{voi.remote for Value of Information calculation - based on Chris Jackson's VoI package}
 %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document is heavily based on the VoI vignette written by Christopher Jackson, and is meant to show the compatiblity of the voi.remote packae with the original voi package. The `voi.remote` package has the exact same functionalities as `voi` package, but allows users to run all the VoI functions on the server rather than running locally.

### Specifying model inputs (same as voi package)

The inputs should be a data frame with one column per parameter and one row per random sample.

```{r}
set.seed(1) 
nsam <- 10000
inputs <- data.frame(p1 = rnorm(nsam, 1, 1), 
                     p2 = rnorm(nsam, 0, 2))
```

### Specifying model outputs

```{r}
outputs_nb <- data.frame(t1 = 0, 
                         t2 = inputs$p1 - inputs$p2)

outputs_cea <- list( 
  e = data.frame(t1 = 0, t2 = inputs$p1), 
  c = data.frame(t1 = 0, t2 = inputs$p2), 
  k = c(1, 2, 3)
)
```

## Expected value of perfect information

#### Computation using random sampling

```{r}
decision_current <- 2
nb_current <- 1
decision_perfect <- ifelse(outputs_nb$t2 < 0, 1, 2)
nb_perfect <- ifelse(decision_perfect == 1, 0, outputs_nb$t2)
(evpi1 <- mean(nb_perfect) - nb_current)
```

```{r}
opp_loss <- nb_perfect - nb_current
mean(opp_loss)
```

#### Using the `voi.remote` package to calculate EVPI

The `voi.remote` package contains a simple function `evpi.remote` to compute the EVPI using the above procedure. The function called the corresponding function in the voi package on the server.

```{r}
library(voi.remote)
```

```{r}
evpi.remote(outputs_nb)
evpi.remote(outputs_cea)
```

#### Analytic computation

```{r}
prob_correct <- 1 - pnorm(0, 1, sqrt(5))
```

```{r}
mean_truncnorm <- function(mu, sig, lower=-Inf, upper=Inf){ 
  a <- (lower-mu)/sig
  b <- (upper-mu)/sig
  mu + sig * (dnorm(a) - dnorm(b)) / (pnorm(b) - pnorm(a))
}
enb_correct <- mean_truncnorm(1, sqrt(5), lower=0) 
mean_nb_perfect <- enb_correct * prob_correct
(evpi_exact <- mean_nb_perfect - nb_current)
```

## Expected value of partial perfect information

The function `evppi.remote` can be used to compute this, similar to the evpi function (which is called on the server).

### Invoking the `evppi.remote` function.

**(a) As a vector**.

```{r}
evppi.remote(outputs_nb, inputs, pars=c("p1","p2"))
```

**(b) As a list**.

```{r}
evppi.remote(outputs_nb, inputs, pars=list("p1",c("p1","p2")))
evppi.remote(outputs_cea, inputs, pars=list("p1",c("p1","p2")))
```

The `evppi.remote` function returns a data frame with columns indicating the parameter (or parameters), and the corresponding EVPPI. If the outputs are in cost-effectiveness analysis format, then a separate column is returned indicating the willingness-to-pay.

```{r}
evppi.remote(outputs_cea, inputs, pars=list("p1",c("p1","p2")))
```

### Changing the default calculation method

#### Gaussian process regression

```{r}
evppi.remote(outputs_nb, inputs, pars="p1", method="gp", nsim=1000)
```

#### Multivariate adaptive regression splines

```{r}
    evppi.remote(outputs_nb, inputs, pars="p1", method="earth")
```

#### INLA method

```{r, eval=FALSE}
evppi.remote(outputs_nb, inputs, pars=c("p1","p2"), method="inla")
```

#### Tuning the generalized additive model method

```{r}
evppi.remote(outputs_nb, inputs, pars=c("p1","p2"), method="gam", gam_formula="s(p1) + s(p2)")

evppi.remote(outputs_nb, inputs, pars="p1", se=TRUE, B=100)
```

#### Single-parameter methods

```{r}
evppi.remote(outputs_nb, inputs, pars="p1", n.blocks=20, method="so")

evppi.remote(outputs_nb, inputs, pars="p1", method="sal")
```

### Traditional Monte Carlo nested loop method

#### Model evaluation function

```{r}
model_fn_nb <- function(p1, p2){ 
  c(0, p1 - p2) 
}
```

```{r}
model_fn_cea <- function(p1, p2){ 
  rbind(e = c(0, p1), 
        c = c(0, p2)) 
}
```

#### Parameter simulation function

```{r}
par_fn <- function(n){
  data.frame(p1 = rnorm(n, 1, 1),
             p2 = rnorm(n, 0, 2))
}
```

#### Invoking `evppi_mc.remote`
```{r}
evppi_mc.remote(model_fn_nb, par_fn, pars="p1", ninner=1000, nouter=100)
```

#### Accounting for parameter correlation
```{r}
par_fn_corr <- function(n, p1=NULL){
  p1_new <- if (is.null(p1)) rnorm(n, 1, 1) else p1
  data.frame(p1 = p1_new,
             p2 = rnorm(n, p1_new, 2))
}
evppi_mc.remote(model_fn_nb, par_fn_corr, pars="p1", ninner=1000, nouter=100)
```
## Expected value of sample information

### Function to generate study data
```{r}
datagen_normal <- function(inputs, n=100, sd=1){
  data.frame(xbar = rnorm(n = nrow(inputs),
                          mean = inputs[,"p1"],
                          sd = sd / sqrt(n)))
}

set.seed(1)
evsi.remote(outputs_nb, inputs, datagen_fn = datagen_normal, n=c(10,100,1000))
```

### Built-in study designs
```{r}
evsi.remote(outputs_nb, inputs, study = "normal_known", n=c(100,1000), pars = "p1")
evsi.remote(outputs_cea, inputs, study = "normal_known", n=c(100,1000), pars = "p1")
```

### Importance sampling method
```{r}
likelihood_normal <- function(Y, inputs, n=100, sig=1){
  mu <- inputs[,"p1"]
  dnorm(Y[,"xbar"], mu, sig/sqrt(n))
}
evsi.remote(outputs_nb, inputs, datagen_fn = datagen_normal, likelihood = likelihood_normal, 
            n=100, pars = "p1", method="is", nsim=1000)
evsi.remote(outputs_nb, inputs, study = "normal_known", n=100, pars = "p1", method="is", nsim=1000)
```
### Further work

## Value of Information in models for estimation

```{r}
p1 <- rbeta(10000, 5, 95)
```

$$p_2 = expit(logit(p_1) + \beta)$$.

```{r}
beta <- rnorm(10000, 0.8, 0.4)
p2 <- plogis(qlogis(p1) + beta)
plot(density(p1), lwd=2, xlim=c(0,1), main="")
lines(density(p2), col="red", lwd=2)
legend("topright", col=c("black","red"), lwd=c(2,2), 
       legend=c("Surveyed infection probability", "True infection probability"))
```

### EVPI and EVPPI for estimation

```{r}
var(p2)
```

```{r}
inputs <- data.frame(p1, beta)
(evppi_beta <- evppivar.remote(p2, inputs, pars="beta"))
(evppi_p1 <- evppivar.remote(p2, inputs, pars="p1"))


sqrt(var(p2)) # or sd(p2)
sqrt(var(p2) - evppi_beta$evppi)
sqrt(var(p2) - evppi_p1$evppi)
```

### How regression-based EVPPI estimation works

```{r}
plot(x=p1, y=p2, pch=".")
mod <- mgcv::gam(p2 ~ te(p1, bs="cr"))
p1fit <- fitted(mod)
lines(sort(p1), p1fit[order(p1)], col="blue")
```

```{r}
p1res <- p2 - p1fit
var(p2) - var(p1res)
```

$$var(Y) - E_{X}\left[var_{Y| X}(Y |X)\right] = var_{X} \left[E_{Y|X}(Y|X)\right]$$
```{r}
var(p1fit)
```

### EVSI for estimation

```{r}
evsivar.remote(p2, inputs, study = "binary", pars="p1", n=c(100,1000,10000))


inputs_p2 = data.frame(p2 = p2)
evsivar.remote(p2, inputs=inputs_p2, study = "binary", pars="p2", n=c(100, 1000, 10000))
```
